Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-12-05 ms/spaceness.tex
final edits(?)

To avoid hard-to-evaluate approximations, we here used forward-time, individual-based simulations
across continuous geographical space.
The question of what regulates real populations has a long history and many answers
\citep[e.g.,][]{lloyd1967crowding,antonovics1980ecological,crawley1990population}, \llabel{ecolcites}
but it is clear that populations must at some point have density-dependent feedback on population size,
or else they would face eventual extinction or explosion.
In the absence of unrealistic global population regulation,
this regulation must be local,
and there are many ways to achieve this \citep{bolker2003spatial}.
In our simulations,
each individual's probability of survival is a decreasing function of local population density,
which shifts reproductive output towards low-density regions,
and produces total census sizes that fluctuate around an equilibrium. \revpoint{3}{2}
This also prevents the population clumping seen by \citet{Felsenstein1975}
(Supplemental \autoref{fig:wm_v_dd})). \revpoint{3}{1}
Such models have been used extensively in ecological modeling
\citep{durrett1994importance,bolker1997using,law2003population,fournier2004microscopic,champer2019suppression}
but rarely in population genetics, \revpoint{AE}{2}
where to our knowledge implementations of continuous space models
before their availability through SLiM \citep{Haller2019}
have focused on a small number of genetic loci
\citep[e.g.,][]{slatkin1989comparison,barton2002neutral,robledoarnuncio2010isolation,rossine2014espao},
which limits the ability to investigate the impacts of continuous space on genome-wide genetic variation as is now routinely sampled from real organisms. 
By simulating chromosome-scale sequence alignments and complete population histories we are able to treat our simulations as real populations and replicate the sampling designs and analyses commonly conducted on real genomic data.

@@

Any simulation of a population of reproducing organisms
requires some kind of control on population sizes,
or else the population will either die out or grow very large after a sufficiently long period of time. \llabel{demog_disco}
The usual choice of population regulation for population genetics --
a constant size, as in the Wright--Fisher model --
implies biologically unrealistic interactions between geographically distant parts of the species range.
Our choice to regulate population size by including a local density-dependent control on mortality
is only one of many possible ways to do this.
We could have instead regulated fecundity, or recruitment, or both;
this general class of models
is sometimes referred to as the ``Bolker--Pacala model'' \citep{bolker1997using}.
It is not currently clear how much different choices of demographic parameters,
or of functional forms for the regulation,
might quantitatively affect our results,
although the general predictions should be robust to similar forms of regulation.
Since populations are still entirely \emph{intrinsically} regulated,
our model still has a very strong ``population genetics'' flavor.
Alternatively, population size could be regulated by interactions with other species
(e.g., a Lotka-Volterra model),
or extrinsically specified by local resource availability
(e.g., by food or nest site availability).
Indeed, our model could be interpreted as a caricature of such a model:
as local density increases, good habitat is increasingly occupied,
pushing individuals into more marginal habitat and increasing their mortality.
Many such models should behave similarly to ours,
but others (especially those with local population cycling), may differ dramatically.
There is a clear need for greater exploration of 
the consequences for population genetics of ecologically realistic population models.

>>>>

To avoid hard-to-evaluate approximations, we here used forward-time, individual-based simulations
across continuous geographical space.
The question of what regulates real populations has a long history and many answers
\citep[e.g.,][]{lloyd1967crowding,antonovics1980ecological,crawley1990population}, \llabel{ecolcites}
but it is clear that populations must at some point have density-dependent feedback on population size,
or else they would face eventual extinction or explosion.
In the absence of unrealistic global population regulation,
this regulation must be local,
and there are many ways to achieve this \citep{bolker2003spatial}.
In our simulations,
each individual's probability of survival is a decreasing function of local population density,
which shifts reproductive output towards low-density regions,
and produces total census sizes that fluctuate around an equilibrium. \revpoint{3}{2}
This also prevents the population clumping seen by \citet{Felsenstein1975}
(Supplemental \autoref{fig:wm_v_dd})). \revpoint{3}{1}
Such models have been used extensively in ecological modeling
\citep{durrett1994importance,bolker1997using,law2003population,fournier2004microscopic,champer2019suppression}
but rarely in population genetics, \revpoint{AE}{2}
where to our knowledge implementations of continuous space models
before their availability through SLiM \citep{Haller2019}
have focused on a small number of genetic loci
\citep[e.g.,][]{slatkin1989comparison,barton2002neutral,robledoarnuncio2010isolation,rossine2014espao,jackson2014landscape},
which limits the ability to investigate the impacts of continuous space on genome-wide genetic variation as is now routinely sampled from real organisms. 
By simulating chromosome-scale sequence alignments and complete population histories we are able to treat our simulations as real populations and replicate the sampling designs and analyses commonly conducted on real genomic data.

@@

Any simulation of a population of reproducing organisms
requires some kind of control on population sizes,
or else the population will either die out or grow very large after a sufficiently long period of time. \llabel{demog_disco}
The usual choice of population regulation for population genetics --
a constant size, as in the Wright--Fisher model --
implies biologically unrealistic interactions between geographically distant parts of the species range.
Our choice to regulate population size by including a local density-dependent control on mortality
is only one of many possible ways to do this.
We could have instead regulated fecundity, or recruitment, or both;
this general class of models
is sometimes referred to as the ``Bolker--Pacala model'' \citep{bolker1997using}.
It is not currently clear how much different choices of demographic parameters,
or of functional forms for the regulation,
might quantitatively affect our results,
although the general predictions should be robust to similar forms of regulation.
Since populations are still entirely \emph{intrinsically} regulated,
our model still has a very strong ``population genetics'' flavor.
Alternatively, population size could be regulated by interactions with other species
(e.g., a Lotka-Volterra model),
or extrinsically specified by local resource availability
(e.g., by food or nest site availability).
Indeed, our model could be interpreted as a caricature of such a model:
as local density increases, good habitat is increasingly occupied,
pushing individuals into more marginal habitat and increasing their mortality. 
Many such models should behave similarly to ours,
but others (especially those with local population cycling), may differ dramatically.

Population genetic simulations often use grids of discrete demes, which are assumed to approximate continuous space. 
However, there are theoretical reasons to expect that increasingly fine grids of discrete demes do not approach the continuous model 
\citep{barton2002neutral}.
If continuous space can be approximated by a limit of discrete models, this should be true regardless of the precise details of the discrete model. 
Although we carefully chose parameters to match our continuous models, we found that some aspects of genetic variation
diverged from the continuous case as the discretization got finer. 
This suggests that these models do not converge in the limit. 
However, many populations may indeed be well-modeled as a series of discrete, 
randomly-mating demes if, for example, suitable habitats are 
patchily distributed across the landscape. 
There is a clear need for greater exploration of 
the consequences for population genetics of ecologically realistic population models.
